### 2019-10-10 usecases.tex
split into usecases

%\IEEEraisesectionheading{\section{Introduction}}
\section{Introduction}
\label{sec:intro}
With a user base of more than one-tenth of the world's population,
spreadsheets are by far the most popular medium for ad-hoc exploration
and analysis of data by end-users~\cite{excel-users}. \saj{Studies show that information workers prefer to operate on their data within spreadsheets while shunning enterprise solutions with more advanced analytical features~\cite{chan1996use,raden2005shedding}. One popular joke among those developing business intelligence applications is that  the ``export to excel'' button is the third-most commonly used button from the menu bar, after OK and Cancel~\cite{bakke2016expressive}.}
Spreadsheets enable users to view, structure, and present data
in an intuitive tabular layout, wherein users can map their data and tasks;
this tabular layout is essential to
the popularity of spreadsheets~\cite{nardi1990spreadsheet}.

Using this tabular layout effectively involves
navigation, \ie
{\em ``the process of viewing and manipulating the computer display to
show another portion of the information space''}~\cite{nav-study}.
Navigation is supported
via two unit operations, scrolling and steering.
\saj{Scrolling is the action of moving displayed text or graphics up, down, or across on a computer screen in order to view different parts of the spreadsheet.}
For example, when analyzing data, users may scroll
to compare data across different screens, or to get
a high-level view of the overall spreadsheet.
\saj{Steering, on the other hand, involves clicking the left mouse button and then dragging the mouse pointer through the spreadsheet to select a specific region.} For example, to issue a formula,
users may steer to select
the subset of the data
to be operated on as an argument within the formula.
Most frequently used spreadsheet formulae
require users to perform steering actions~\cite{bradbard2014spreadsheet, lawson2009comparison}.
Overall, both scrolling and steering are crucial as
users navigate spreadsheets to
identify, compare, and summarize data.

\begin{figure*}
     \centering %trim=left bottom right top
  %\includegraphics[width=\linewidth,trim={0 95 0 80},clip]{images/navigation.pdf}
  \includegraphics[width=\linewidth,trim={0 95 0 80},clip]{images/screenshot.pdf}
    \caption{\noah: navigation interface consisting of (a) a zoomable overview and (b) an aggregate column integrated with (c) a spreadsheet. A context bar consisting of (d) a navigation history displaying locations visited so far using the overview, and (e) a breadcrumb showing the current navigation path (\eg Home). (f) The user’s current focus in the spreadsheet is highlighted on the overview. (g) Columns corresponding to the navigation attribute (city) and aggregate column (availability) are highlighted on the spreadsheet.}
  \label{fig:ux}
\end{figure*}

However, navigating spreadsheet data using scrolling or steering
is often challenging, even for moderately sized spreadsheets that only span
a few screens~\cite{network-context,nardi1990spreadsheet}, due to multiple inter-related reasons:
\squishlist
\item {\em Loss of overview and context.}
When navigating
spreadsheets, users can easily lose the context
of where they are and where they should go next~\cite{network-context}.
The only navigational context provided by spreadsheets
is the built-in scrollbar that acts as a one-dimensional
overview and indicates the user's current location on the sheet.
However, since this overview does not capture the
layout and structure of the data,
users are forced to mentally assimilate
the layout and recall it on-demand, as they navigate a spreadsheet.

\item {\em Cognitive and mechanical burdens.}
The lack of contextual cues leads to severe
cognitive and mechanical burdens
for users~\cite{cockburn2009review}.
Users often end up taking their own drastic measures to
avoid getting lost; for example,
some users create personalized overviews
extrinsic to the spreadsheet,
by sketching maps
of spreadsheets on paper~\cite{network-context}.
Other users add their own landmarks such as headers or
colored cells, as a visual affordance to assist
in navigation~\cite{network-context}.
Steering via dragging the mouse pointer across multiple screens to select
a subset of data as input to a formula
can often be challenging as well:
the only remedy is for users to abandon steering entirely and
instead remember the range of
the subset of data of interest, and then correctly enter this range
as the argument to the formula, often giving rise to errors that are increasingly prevalent in spreadsheets~\cite{panko1998we}.

\item {\em Visual discontinuities.}
The limited viewport afforded to the user
introduces a visual discontinuity between the
information being displayed.
For example, comparing spatially separated subsets
of data within the spreadsheet
requires moving back and forth between multiple viewports,
which can be overwhelming~\cite{nardi1990spreadsheet,network-context}.
As an alternative, users tend to copy subsets of data side by side to
reduce the visual discontinuity~\cite{nardi1990spreadsheet,network-context},
which is cumbersome.
\squishend

Overall,  while navigating present-day spreadsheets, {\bf \em users often lose context,
get overwhelmed, and experience visual discontinuities.} 
Addressing these challenges requires considerable manual effort.
As we argue in Section~\ref{sec:related},
existing spreadsheet functionalities such as
pivot tables, named ranges, and subtotals,
while providing partial support for some of these challenges
(\eg pivot table generates a summary
while losing the mapping between the
raw data and the summary; named ranges requires users
to manually associate names with ranges of data),
do not eliminate them entirely.
With the ease of data generation, and with spreadsheets
now supporting data
of increasingly larger scales---\eg Google Sheets
now supports two million cells, a five-fold increase
from the previous limit of 400K cells---navigating
data within spreadsheets will become harder, not easier.

So, how do we support more effective navigation of data
within spreadsheets?
One approach would be to try to integrate
an overview of the overall structure of the data along
with the spreadsheet~\cite{grudin2001partitioning}.
The resulting integration would then be a classical
overview+detail interface
where the spreadsheet is the detailed view.
Overview+detail interfaces are used
to facilitate navigation in various domains such as 
text editors and maps~\cite{cockburn2009review},
and enable users to manipulate
the overview or detailed view,
to perform high-level or low-level operations,
respectively.
Overview+detail interfaces
have been shown to be effective in these domains,
reducing cognitive load for users by
providing the big picture first,
helping them quickly assimilate the information space~\cite{cockburn2009review}.
In this paper, our goal is to extend
existing spreadsheets
with an overview plug-in, giving an overview+detail interface.

However, while an overview plug-in for spreadsheets
does seem appealing and natural,
developing it leads to several challenges.
\squishlist
\item {\em Overview modality.}
One approach for an overview is to simply add a
zoomed out version of the entire sheet as a pane on the side,
as in popular presentation software applications like Microsoft PowerPoint or text editors
like Sublime Text. 
Unfortunately, this approach
would not suffice for a spreadsheet---while it would help users make
somewhat more informed
scrolling decisions, the zoomed out presentation of the overview would be unreadable.
Another approach, adopted by map tools like the
early versions of Google maps,
is to use the overview
to provide a \saj{global} context of the user's current location
\saj{being displayed} in the \saj{zoomed in} detailed view~\cite{cockburn2009review}.
Users can perform zooming operations on the detailed
view as the overview remains static.
Since spreadsheets already d  isplay the raw data,
zooming into and out of a detailed view consisting of
this raw data is not meaningful.
Instead, the goal of our spreadsheet
overview is to provide a big picture view of the spreadsheet,
capturing the
data distribution and \saj{augmenting it with summarization},
while enabling meaningful interactions
for navigating the data.
This necessitates
{\em semantic zooming}~\cite{perlin1993pad} instead of geometric
zooming (unlike text editors or presentation software),
on the overview view instead of the detailed view
(unlike map tools).
Several domains, \eg online maps and program visualization,
adopt semantic zooming
to ensure improved readability and
comprehension of the information displayed~\cite{summers2003experimental}.
However, to the best of our knowledge, no prior work
has ever developed semantic zooming for spreadsheet data.

\item {\em Construction of the overview.}
Given a spreadsheet
with many rows,
it is not clear how to employ semantic zooming
to construct the overview.
How do we group rows together in a ``meaningful''
way
such that this grouping applies to all 
data types, including strings and numbers?
Since we want to allow users to control how the overview
is constructed, we can let them select the grouping
attribute(s).
Following this, we need mechanisms to automatically group nearby rows
into meaningful bins.
Further, the bins need to be constructed
at various granularities,
allowing users to zoom in and zoom out of the overview.
We also need to allow the users
to customize the bins
to construct meaningful categories, while providing
interpretable and intelligent defaults.
Customizing these bins will also eliminate
visual discontinuities: users can simply arrange the bins
the way they want and keep them near each other
for easy comparison.

\item {\em Operations on the overview.}
While the scrolling issues are partially addressed
by having the overview, the steering issues remain unaddressed.
How do we leverage the overview to allow users
to select regions of interest for issuing formulae
without steering
across multiple screens?
As the granularity of the overview changes (during zoom in/out)
how do we ensure that users don't have to reissue the formula from scratch?
For this, we take inspiration from aggregation
in relational databases. In relational databases, each ``group'' is associated
with one or more aggregates. As long as the groups are
appropriately constructed (using the mechanisms alluded to above),
formulae can be issued once without requiring any steering, and
automatically maintained for all groups when
zooming in and zooming out.

\item {\em Seamless integration as a plug-in.}
Finally, the design of the overview
should be generic so that it can be integrated
with any existing spreadsheet tool on any spreadsheet
dataset without impacting
existing functionalities or look-and-feel.
Any new interactions introduced should be consistent
with traditional spreadsheet semantics,
and complement existing spreadsheet interactions.
Interactions between the overview and the raw spreadsheet
should be coordinated to the extent possible.

\squishend

\noindent
We address the aforementioned challenges
in our tool \noah,
an in-situ
\textbf{\underline{n}}avigation interface
for \textbf{\underline{o}}verviewing
and \textbf{\underline{a}}nalyzing
spreadsheet data \textbf{\underline{h}}olistically.
\noah is constructed as a plugin to
an existing spreadsheet tool,
{\scshape DataSpread}~\cite{dataspread},
an open-source scalable web-based spreadsheet.
While \noah's design is not tied to {\sc DataSpread},
we opted not to use other popular spreadsheet
tools like Google Sheets and Microsoft Excel because they
are closed source.
Figure~\ref{fig:ux} shows a snapshot of \noah.
When the user chooses to explore
the data by a specific attribute,
an overview is constructed and displayed
within \noah, next to the raw spreadsheet data (Figure~\ref{fig:ux}a).
The overview is a histogram
and depicts the overall data distribution
for the chosen attribute,
enabling users to assimilate the data.
\noah ensures that there is coordination between
the overview and the spreadsheet:
for example, panning and zooming on the overview
are reflected on the spreadsheet by displaying
the spreadsheet data corresponding to the bin
currently in focus in the overview.
Cumbersome scrolling operations are eliminated in favor
of a few clicks on the overview interface.
% \noah leverages {\scshape DataSpread}'s
% underlying spatial mapping of data to ensure coordination between the overview and the spreadsheet so that users can relate the interactions on the overview with the raw data.
Instead of steering to analyze the data,
users can issue aggregation operations
on the overview,
and view results on a separate {\em aggregate column},
alongside the overview (Figure~\ref{fig:ux}b).
In this manner,
users can issue formulae on different
subsets of the data while remaining on the same screen,
reducing visual discontinuity.
Finally, \noah automatically creates
contextual and historical information
(Figure~\ref{fig:ux}d and~\ref{fig:ux}e)
while displaying visual cues
(Figure~\ref{fig:ux}f and~\ref{fig:ux}g)
so that users don't lose context during navigation.

The primary contribution of our work is twofold:
\squishlist
\item We formalize the design of a general
navigation (overview+detail) interface
for exploration and analysis of large spreadsheets.
We realize this design in the form of \noah,
a plugin to an existing web-based spreadsheet tool,
ensuring that interactions supported by \noah
complement existing spreadsheet operations.
\item We conduct a user study to evaluate
the benefits and limitations of this plugin.
\saj{The study shows that compared to Microsoft Excel, participants were able to complete spreadsheet
tasks involving navigation 
correctly and quickly in \noah. Participants made {\bf $2.5\times$ fewer} mistakes while being {\bf $2\times$ faster} with a \noah-integrated spreadsheet than with Excel.} 
\squishend

\toappendix{The rest of the paper is organized as follows: In Section~\ref{sec:usage} we first present a case study of \noah being applied to a real-world scenario.  
In Section~\ref{sec:related}, we discuss related work. We then explain the design considerations for developing \noah in Section~\ref{sec:design}. In Section~\ref{sec:ui}, we present \noah’s user interface and supported operations. We then discuss the design and outcomes of a user study evaluating the effectiveness of \noah in Section~\ref{sec:study}. Finally, we discuss the limitations of \noah and possible future enhancements in Section~\ref{sec:discuss}.}

>>>>

%!TEX root = main.tex
